---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

Download the dataset from:

https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

and save it into your working directory. Then, read the dataset using the following code:

```{r}
act <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
act1<-act[!is.na(act$steps),]
act1<-aggregate(steps ~ date, data=act1 ,sum)
hist(act1$steps,main = "Histogram of Steps vs Date" ,xlab= "Number of Steps" ,col = "pink")
abline(v = mean(act1$steps), col="red")
abline(v = median(act1$steps), col="blue")
```

To calculate the mean and median:
```{r}
mean(act1$steps)
median(act1$steps)
```
It follows that mean = 10766.19 and median = 10765.


## What is the average daily activity pattern?

We can plot the graph of the dayily activity pattern using the following code:
```{r}
act1<-act[!is.na(act$steps),]
act2<- aggregate(steps ~ interval, data = act1, mean)
with(act2, plot(interval, steps, type="l", xlab = "Intervals", ylab = "Average Number of Steps"))
```

The maximum average step across intervals is 206.1698:

```{r}
max(act2$steps)
```
And in happens in the interval=835:

```{r}
act2[act2$steps>206,]
```

## Imputing missing values
Total number of missing values:

```{r}
act1<-act[!is.na(act$steps),]
NumOfMissValues<- nrow(act)-nrow(act1)
NumOfMissValues
```

I will use the median of steps per day (as was calculated above) to fill the NAs.

```{r}
act3<-tapply(act$steps, act$date, sum)
act3<-as.data.frame(act3)
act3[is.na(act3)]<- 10765
colnames(act3) <- "steps"
head(act3)
```

The new histogram will be:

```{r}
hist(act3$steps, col = "pink")
abline(v = mean(act3$steps), col="red")
abline(v = median(act3$steps), col="blue")
```

Although the total amount of days with 10000-15000 steps is increased (because we just replaced all the NA values with 10765 which falls in this interval), the new median and mean are almost the same as before:

```{r}
mean(act3$steps)
```

```{r}
median(act3$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
act4<-cbind(date = row.names(act3), act3)
act4<-cbind(act4, weekday=weekdays(as.Date(act4$date)))
head(act4)
```

```{r}
act4$weekday<-as.character(act4$weekday)
act4$weekday[act4$weekday %in% c("Saturday" , "Sunday")]<- "weekend"
act4$weekday[!(act4$weekday %in% c("weekend"))]<- "weekday"

head(act4, n = 14)
```

```{r}
act5<-cbind(act1, weekday=weekdays(as.Date(act1$date)))
act5$weekday<-as.character(act5$weekday)
act5$weekday[act5$weekday %in% c("Saturday" , "Sunday")]<- "weekend"
act5$weekday[!(act5$weekday %in% c("weekend"))]<- "weekday"

act5<- aggregate(steps ~ interval + weekday, data = act5 , mean)

par(mfrow= c(2,1))
act5_weekday <- subset(act5, act5$weekday == "weekday")
act5_weekend <- subset(act5, act5$weekday == "weekend")
plot(act5_weekday$interval, act5_weekday$steps, type = "l" , main = "Weekdays", xlab = "Intervals", ylab = "Steps")
plot(act5_weekend$interval, act5_weekend$steps, type = "l", main = "Weekends", xlab = "Intervals", ylab = "Steps")


```


